---
title: mb-article HTML Format for Quarto
subtitle: A 3-column academic format for Quarto posts and articles
date: 2025-11-12
date-modified: last-modified
draft: false
freeze: false
thanks: Special acknowledgments and funding sources
doi: "123.456"
bibliography: biblio.bib
author:
  - name: Author-1
    linkedin: mbacou 
    email: author-1@mail.com
    orcid: 0000-0003-1810-6820
    affiliation: My Affiliation 1    
  - name: Author-2
    email: author-2@mail.com
    orcid: 0000-0003-1810-6819
    affiliation: My Affiliation 2
categories:
  - cat-1
  - cat-2
  - cat-3
keywords: 
  - keyword 1
  - keyword 2
  - keyword 3

abstract: |
  Abstract text -- Lorem Ipsum is simply dummy text of the printing and typesetting industry. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. 
  It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.

description: |
  **mb-article** is a Quarto extension for HTML formats with an opinionated academic design and simplified navigational elements.

format: mb-article-html
title-block-banner: url(banner.avif)
---

# Introduction

This document is an example that demonstrates the formatting and features of the current HTML format extension for Quarto.

Here is a sample margin note. The left column can also include side margins, but note that they will display above the TOC and branding (see below).

::: {.aside .text-serif .fs-1 .mb-2}
aside -- A note placed in the page margin with CSS classes `.aside .text-serif`.
:::

Lorem Ipsum is simply dummy text of the printing and typesetting industry. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book.

::: {.aside .bg-danger-subtle .text-danger .p-2}
aside -- A note placed in the page margin with CSS class `.aside .bg-danger-subtle .text-danger`.
:::

Lorem Ipsum is simply dummy text of the printing and typesetting industry. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book.

**More Information**

You can learn more about controlling the appearance of HTML output at <https://quarto.org/docs/output-formats/html-basics.html>.

# Brand Colors

Bootstrap colors inherit from [_brand.yml](_brand.yml) so customizations are straightforward.

<div class="d-flex gap-1 text-sans mb-2">
<div class="flex-fill p-1 text-bg-info">Info</div>
<div class="flex-fill p-1 text-bg-success">Success</div>
<div class="flex-fill p-1 text-bg-warning">Warning</div>
<div class="flex-fill p-1 text-bg-danger">Danger</div>
</div>
<div class="d-flex gap-1 text-sans">
<div class="flex-fill p-1 text-bg-primary">Primary</div>
<div class="flex-fill p-1 text-bg-secondary">Secondary</div>
<div class="flex-fill p-1 text-bg-light">Light</div>
<div class="flex-fill p-1 text-bg-dark">Dark</div>
</div>

# Typography

::: {.column-sidebar .bg-info-subtle .p-2 .text-info}
column-sidebar -- A note placed in the left column with CSS class `column-sidebar`.
:::

Bootstrap default line-height and font sizing typically work, but this format enforces a strict vertical rhythm whenever possible[^1]. I made sure to define all vertical spacing and padding as multiples of a standard line height of `$spacer`.

[^1]: About **vertical rhythm** and why it matters for content readability, see [Klaas Leussink](https://hnldesign.hashnode.dev/setting-a-flexible-baseline-grid-in-css) article on his personal blog.

No additional changes to font definitions aside from using `Crimson Pro` and `M+ Code` fonts[^2] by default, but always customizable via `_brand.yml` utility.

[^2]: M+ font system was designed by Japanese artist [Coji Morishita](https://mplusfonts.github.io/) with extended support for Latin, Japanese, Cyrillic. It is an excellent monospace font for smaller display sizes.

::: {.aside}
![](logo.svg){.rounded-circle}

A rounded image with `.rounded-circle` moved to the margin.
:::

Here are **bold**, *italic*, and <small>smaller</small> text spans. Here are ^superscript^ and ~subscript~ characters, and ~~strikethrough~~ notes. Standard Bootstrap semantic colors with e.g. [.text-info]{.text-info} provide accented text. In addition this theme uses [.text-muted]{.text-muted} to de-emphasize content such as post metadata, margin notes, and captioning.

## Heading 2

> Quoted lines.
> <small>With a signature</small>

### Heading 3

Sample paragraph content with list of items:

-   List 1
    -   List 1.1
    -   List 1.2
-   List 2
-   List 3

Ordered list:

1.  List 1
2.  List 2
3.  List 3

Finally a normal rule `hr` divider.

------------------------------------------------------------------------

And a paragraph of content after the rule.

#### Heading 4

By default citations (and cross-references) are read in from an external `references.bib` file, but this may be edited in the YAML front matter as usual. Each citation must have a key, composed of ‘\@’ + the citation identifier from the database. Here is a margin reference to [@xie2015].

##### Heading 5

Here is a long footnote reference[^3]. Caption and footnote placement can be controlled in the config files `_quarto_yml` and/or `_metadata.yml`, or directly in the YAML front matter.

[^3]: Here's a margin note with multiple paragraph blocks and a fancy section header.

    ##### Heading 5

    With subsequent paragraphs indented to show that they belong to the previous footnote.

###### Section 6

Headers at level 6 are used as fancy section dividers.


# Figures & Plots

```{r}
#| column: page-right
#| fig-cap: A sample dynamic graph

library(ggplot2)

ggplot(mtcars, aes(factor(carb), mpg, fill = factor(carb))) +
  geom_col() +
  guides(fill = guide_legend(nrow = 1)) +
  labs(
    title = "Custom ggplot with Bootstrap brand",
    subtitle = "Theme provided by R package `mblabs`",
    caption = "A long plot caption with many references
    over multiples lines."
  ) +
  theme(legend.position = "bottom")
```


# Tables

I prefer my tables to resemble LaTeX typesetting, so a few simple changes here.

|        |      |
|--------|-----:|
| apple  | 2.05 |
| pear   | 1.37 |
| orange | 3.09 |

: Caption: my default table spacing

| fruit  | price |
|--------|-------|
| apple  | 2.05  |
| pear   | 1.37  |
| orange | 3.09  |

: Caption: compact spacing borderless {.borderless .hover .sm}

| fruit  | price |
|--------|-------|
| apple  | 2.05  |
| pear   | 1.37  |
| orange | 3.09  |

: Caption: striped hover {.striped .hover}

::: column-page-right
A table running over the right margin.

| fruit  | price |
|--------|-------|
| apple  | 2.05  |
| pear   | 1.37  |
| orange | 3.09  |

: Caption: striped hover secondary {.striped .hover .secondary}
:::

::: column-page-left
A table running over the left sidebar.

| fruit  | price |
|--------|-------|
| apple  | 2.05  |
| pear   | 1.37  |
| orange | 3.09  |

: Caption: striped primary {.striped .primary}
:::

```{r}
#| column: page
#| tbl-cap: A full-width table to increase content space

knitr::kable(mtcars[1:6, 1:10])
```





